Rusty Linux: Advances in Rust for Linux Kernel Development

### 2024-07-25 paper.tex
Update on Overleaf.

%\documentclass[sigconf]{acmart}
\documentclass[sigconf,review,anonymous]{acmart}
\usepackage{listings, listings-rust}
\usepackage{xcolor}
\usepackage[raggedrightboxes]{ragged2e}

@@

%% \author{Shane Panter}
%%     \affiliation{%
%%     \institution{Boise State University}
%%     \city{Boise}
%%     \state{Idaho}
%%     \country{USA}}
%%     \email{shanepanter@boisestate.edu}

%% \author{Nasir U. Eisty}
%%     \affiliation{%
%%     \institution{Boise State University}
%%     \city{Boise}
%%     \state{Idaho}
%%     \country{USA}}
%%     \email{nasireisty@boisestate.edu}

@@

\received{30 May 2024}
\received[revised]{TBD}
\received[accepted]{TBD}

@@

The 1995 movie Hackers~\cite{Wikipedia_contributors2024-zr} prescient predictions regarding the ease
of breaking into computing systems in cyberspace have come to fruition. The White House Office of
the National Cyber Director (ONCD) released a report calling for the technical community to
proactively reduce the attack surface in cyberspace with a two-pronged approach~\cite{United_States_Gov2024-pp}. First, we need to
address the root cause of many of the most heinous cyber attacks: memory-unsafe programming
languages. Second, we need to establish better cybersecurity quality
metrics so we can have a better understanding of the cybersecurity landscape.

@@

\subsubsection{\textbf{Research Questions}}
\label{sec:researchQuestions}

We have defined the following research questions to drive this study.

\begin{itemize}
    \item \textbf{RQ1:} What are the existing approaches for implementing operating system kernels in Rust?

    All though Rust in the Linux kernel is very recent we would like to see what approaches researches have explored in any operating system. Experiential and research kernels can provide new ideas that can be integrated into the Linux kernel.
    \item \textbf{RQ2:} What are the performance implications of using Rust for operating system kernel development?

    The C programming language has the advantage of 40 years of improvements and lessons learned regarding how to write fast code. Rust being a new language still has many unexplored areas with how to write fast code especially in kernel space.
      %in terms of throughput, latency, and resource utilization?
    \item \textbf{RQ3:} What are the major challenges and limitations when
      developing operating system kernels in Rust?% particularly in comparison to other languages?

      Kernel space presents unique challenges for the software engineer using Rust. Rust was not designed to implement kernels and is more restrictive than C. Thus, this research question is looking for any stumbling blocks that have already been found.
    \item \textbf{RQ4:} What are the lessons learned when developing operating systems kernels in Rust?

    After the work has been completed we want to find all the research that reflects on what went well and what didn't.
\end{itemize}

@@

We employed a robust multi-step search strategy across three databases: ACM Digital Library,
IEEE Xplore, and Scopus to find all the current research. 
%These databases cover a wide range of current software engineering and computer science literature. 
We leveraged the advanced search features of all three databases to search the titles and abstracts for
keywords using boolean search operators.

@@

We searched all three databases with the keywords listed in table~\ref{tab:keywords} between January
1, 2019 and April 1, 2024. While both ACM and IEEE had mutually exclusive results, Scopus had 44
duplicates that needed to be removed. We then used Google Scholar for forward and backward
snowballing. We choose to look at only the previous 5 years of research in order to capture the
bleeding-edge research that is currently being done.

\begin{table}
\begin{tabular}{|| p{2cm}| p{4cm} | p{1cm} ||}
 \hline
 Database & Query & Results \\
 \hline\hline
 ACM  & (Abstract:("operating system" kernel linux OS) AND Abstract:(rust) )OR (Title:("operating
 system" OS kernel linux) AND Title:(rust))  & 30 \\
 IEEE & ("All Metadata":"operating system" AND "All Metadata":rust) OR ("All Metadata":kernel AND
 "All Metadata":rust) OR ("All Metadata":linux AND "All Metadata":rust) & 39 \\
 Scopus & TITLE-ABS-KEY ( ( "operating system" OR kernel OR linux ) AND rust ) AND PUBYEAR > 2018
 AND PUBYEAR < 2025 AND ( LIMIT-TO ( SUBJAREA , "COMP" ) ) AND ( LIMIT-TO ( LANGUAGE , "English" ) )
 & 87 \\
 \hline
\end{tabular}
\caption{Search Queries used for each Database.}
\label{tab:keywords}
\end{table}

@@

Following the guidelines outlined by Kitchenham and Charters~\cite{Stuart2007-cc} we set the
inclusion and exclusion criteria based on our research questions outlined in
section~\ref{sec:researchQuestions}. We only consider papers that are written in English, published
in conferences, journals, workshops, and within the time frame of 2019 and 2024. The published papers should
describe using the Rust programming language for either developing a new kernel, extending an
existing kernel, or authoring drivers. We include any type of kernel architecture, including
monolithic kernel, microkernel, or unikernel in both the embedded and non-embedded space, as long as
the paper is using the Rust programming language in some way. Papers that describe solutions that
reside 100\% in user space will be excluded. Early efforts to use Rust in kernel development
were impacted by changes and updates in both the Rust language and compiler and relied on unstable
features, which are thus excluded from this review.

@@

Once we had our initial collection of papers from the databases, we merged all duplicate papers into
one record and then started our first pass, which consisted of reading the title and abstract and
either marking the paper as \textit{include} or \textit{exclude}. Once the initial screening was
completed we were left with 26 potential papers that we needed an in-depth reading with an
additional 6 papers were found during forward and backward snowballing.

@@

For the first approach, Boos et al.~\cite{Boos2020-zh} created an experimental operating system
named Theseus that operates in a single address space and single privilege level and uses properties
of the Rust programming language to realize isolation instead of relying on hardware. This novel
approach pushes some of the roles that a traditional operating system would take onto the
compiler. The design of Theseus uses a novel cell-based structure where ownership of memory and
resources is enforced by the compiler, thus avoiding these checks during runtime to get improved
performance. In addition to memory safety, Theseus has a goal of addressing the state spill
problem. State spill happens when a single service in an operating system can harbor a state change
induced by interacting with other services. That state change can eventually cause a system crash or
system instability at some point in the future making it very difficult to track down the problem
due to the separation between when the problem occurred and when it was observed. Theseus OS is an
operating system that was designed to match the Rust language instead of the more traditional route
of matching the hardware. Li and Sato~\cite{Li2024-yb} explored using Rust to write an exokernel named W-Kernel. The authors proposed a novel architecture for an OS by embedding a WebAssembly(WASM) runtime into the kernel that can run programs written in any language that can compile down to WASM.

For the second approach researches take a more  traditional micro-kernel implementation~\cite{Chen2023-wb, Liang2021-bo}. By pushing most kernel functionality
to user space, the authors limited the surface area that needs to be proved. Chen X et al.~\cite{Chen2023-wb} took
their implementation one step further by formally verifying their new micro-kernel named Atmosphere
by using both the liner type system of Rust in combination with a SMT solver. They were
able to get a 7.5:1 proof-to-code ratio, which is higher than other formerly verified micro-kernels
SeL4 and CeriKOS both have proof code ratios of 19:1 and 20:1, respectively. While this
approach was novel drivers running in userspace are still not formally verified and don't have to be
written in Rust.

Finally researchers have taken existing Unikernels written in C and ported them to to Rust. Unikernels are designed to do one thing
and one thing only, thus limiting the complexity that a general-purpose operating system has to
deal with~\cite{Madhavapeddy2014-zw}. This reduced complexity allowed  Lankes et al.~\cite{Lankes2019-cm} to take HermitCore and rewrite it in Rust. The new
Unikernel named RustyHermit consists of only 3.27\% unsafe Rust, with the rest of the code base consisting of safe Rust, dramatically decreasing the surface area where memory errors could originate.

@@

@@

Culic et al.~\cite{Culic2022-bk} looked at latency issues in Tock. Tock is a new operating
system written in Rust that is designed to run on embedded systems but does not provide Real-time capabilities. The
authors attempted to add real-time capabilities by integrating eBPF into the Tock kernel to improve
the interrupt handler's response time. The Authors found that early work (still in the prototype
stage) lowers the response times of the system and interrupts response times 3x.

@@

Li et al.~\cite{Li2024-be} explored the feasibility of using Rust in kernel space. The authors took
an existing component, the Out of Memory (OOM) and implemented a replacement using the Rust
programming language. The non-encapsulated interface Rust component which was almost identical to
the original C component, only introduced a 0.7\% overhead. The encapsulated Rust component, on the
other hand, added a 3\% performance overhead. Another important aspect of the kernel is size, and we must
be careful to keep the size of the Rust component as close as we can to the C component so we can
still run on the same hardware. The authors found that there was only a 0.06\% size increase
when compared to the original C implementation.

\begin{table}
    \begin{tabular}{||l|l|l||}
    \hline
    No. & Implication & Studies that Reported the challenge\\
    \hline\hline
    1 & Performance & \cite{Gonzalez2023-ek}, \cite{Li2024-be}, \cite{Ma2023-ef}\\
    2 & Throughput & \cite{Gonzalez2023-ek}\\
    3 & Latency & \cite{Culic2022-bk} \\
    \hline
  \end{tabular}
  \caption{Performance Implications of using Rust in the Kernel}
    \label{tab:RQ2}
\end{table}

@@

\begin{table*}[htbp]
    \begin{tabular}{p{1cm} p{5cm} p{10cm}}
        \hline
        No. & Challenge & Description\\
        \hline
        1 & Binary Size~\cite{Ayers2022-sf}  &
        \begin{itemize}
            \item Deeply ingrained monomorphization which increased the size of the rust binaries
            \item Compiler optimizations are not as mature as some C based compilers, thus increasing binary size
            \item Hidden data structures and data
            \item Sub-optimal compiler generated support code
        \end{itemize}
        \\
        \hline
        2 & Missing Features~\cite{Burtsev2021-mh} &
        \begin{itemize}
            \item Supporting trait bounds on functions and closures with any number of arguments
            \item Expose type information in procedural macros
            \item Support a collision-free, unique type identifier

            \item Support typed assembly language for Rust
            \item Support trusted build environments
            \item Provide software-only stack guard with extensible probing interface
            \item Develop zero-copy serialization of “plain- old” data structures
        \end{itemize}
        \\
        \hline
        3 & Soundness~\cite{Klimt2023-ob} &
        \begin{itemize}
            \item Unsynchronized Global State - any use of mutable statics is unsafe
            \item C-Style Abstractions - Use rust style abstractions to properly encapsulate internal unsafe usage of raw pointers
            \item Aliasing Mutable References - Giving out raw pointers to memory that is also referenced mutably
            \item Re-implementing Memory Access - instead of accessing specific memory regions in assembly, creating and using references to the whole region is preferable
        \end{itemize}
        \\
        \hline
        4 & Panics \cite{Ma2023-ef},\cite{Burtsev2021-mh} &
        \begin{itemize}
            \item Support extendable, no\_std unwind library
            \item Stack unwinding in embedded environments
        \end{itemize}
        \\
        \hline
        5 & C Interop \cite{Miller2019-xm}, \cite{Li2021-xo} &
        \begin{itemize}
            \item Kernel interfaces, while designed for extensibility, are not designed for type safety
            \item Hybrid Code Flow. The Rust compiler can not track ownership when switching between modules written in C and Rust
        \end{itemize}
        \\
    \hline
  \end{tabular}
  \caption{Challenges Unique to the Rust Programming Language}
    \label{tab:RQ3}
\end{table*}

@@

A primary concern for programming languages that are used for operating system development is the
size of the binary especially in an embedded environment.  Ayers et al.~\cite{Ayers2022-sf} focused
on reducing the size of binaries produced by the rust compiler while working with Tock OS. They were
able to identify several  causes of binary growth that are specific to the Rust programming language
and have the following 5 recommendations when using Rust in a size-constrained environment:
\begin{itemize}
  \item Minimize Length + Instantiations of Generic Code
  \item Use Trait Objects Sparingly
  \item Don't Panic
  \item Carefully Use Compiler Generated Support Code
  \item Don't use static mut
\end{itemize}

@@

Klimt et al.~\cite{Klimt2023-ob} details the lessons learned and challenges when implementing
Theseus. First they found that it is impossible to write a complete operating system in 100\% safe
Rust. For example, when writing a memory management system, raw pointers must be used to
modify the hardware. Despite the limitations imposed by the hardware, the authors describe how
Theseus leveraged intralingual design to maximize the compiler's role in enforcing correctness. By
leveraging Rust's type system and borrow checker memory safety, correct ownership transfer can be
achieved at a higher level than what could be done in C. The authors also detailed some of the
limitations of intralingual design, such as not being as expressive as many other formal verification
techniques due to the limited invariance that the type system can enforce. One of the most
important lessons learned was the insight that a linear type system itself cannot guarantee
the uniqueness of the resource represented, such as when a memory resource may overlap. The authors
introduce the idea of using a hybrid approach of verification where they leverage both the linear
type system and an SMT solver. The authors also explore the bootstrapping problem with rust
systems. In operating systems written in C, the kernel provides the ownership root to applications running on
top. Finding a new ownership root is an open research question with regard to using Rust for
operating system development.

@@

\textit{\textbf{Internal Threats.}} Internal validity refers to the degree to which the review accurately establishes
causal relationships. Potential threats include:

\begin{itemize}
  \item Publication Bias: Studies demonstrating positive outcomes of using Rust for kernel
    development may be more likely to be published and cited, leading to an overestimation of its
    benefits.
  \item Quality of Studies: The methodological rigor of the included studies varies, and some may
    suffer from design flaws or biases that were not adequately controlled for, thus affecting the
    reliability of their findings.
  \item  Selection Criteria: The inclusion and exclusion criteria for selecting relevant studies might inadvertently bias the review         towards certain types of research, such as those reporting successful integration while
    underrepresenting studies detailing challenges and failures.
\end{itemize}

\textit{\textbf{External Threats.}} External validity pertains to the generalizability of the review's findings
beyond the specific context of the reviewed studies. Threats to external validity include:

\begin{itemize}
  \item Context-Specific Findings: Many studies focus on specific kernel modules or use-cases, which
    may not be representative of the broader kernel environment. The success of Rust in isolated
    components does not necessarily translate to the entire kernel.
  \item Temporal Changes: The rapidly evolving nature of both the Rust programming language and the
    Linux kernel means that findings from older studies may no longer be applicable, as improvements
    and changes in both domains can alter the landscape significantly.
    \item Definition Ambiguity: The term "Kernel development in Rust" encompasses a wide range of activities,
    from minor module development to complete subsystem rewrites. Variations in how researchers and
    developers interpret this integration could lead to inconsistent findings.
\end{itemize}

\textbf{\textit{Conclusion Threats}.} Conclusion validity involves the degree to which the conclusions drawn from the
review are credible and supported by the data. Potential threats include:

\begin{itemize}
  \item Heterogeneity of Studies: The diverse methodologies, metrics, and contexts of the included studies
    can lead to challenges in synthesizing findings and drawing unified conclusions. For example, it
    is very difficult to directly compare a micro-kernel, Unikernel, and hybrid Kernel.
  \item Reviewer Bias: Personal biases of the reviewers in interpreting data and making judgments
    about study quality and relevance may skew the results
\end{itemize}

Efforts were made to minimize this through predefined criteria and multiple reviewers, but some
degree of subjectivity is inevitable.  In addressing these threats, we employed rigorous methods for
study selection, data extraction, and analysis, and we remain transparent about the
limitations. Future reviews could mitigate these threats further by incorporating a broader range of
sources, using more precise definitions, and employing meta-analytic techniques to quantitatively
synthesize findings. Despite these threats, our review provides a valuable synthesis of the current
state of research on the use of Rust in kernel development, highlighting both its potential benefits
and the challenges that need to be addressed.

>>>>

\documentclass[sigconf]{acmart}
%\documentclass[sigconf,review,anonymous]{acmart}
\usepackage{listings, listings-rust}
\usepackage{xcolor}
\usepackage[raggedrightboxes]{ragged2e}

@@

\author{Shane K. Panter}
    \affiliation{%
    \institution{Department of Computer Science}
    \institution{Boise State University}
    \city{Boise}
    \state{Idaho}
    \country{USA}}
    \email{shanepanter@boisestate.edu}

\author{Nasir U. Eisty}
    \affiliation{%
    \institution{Department of Computer Science}
    \institution{Boise State University}
    \city{Boise}
    \state{Idaho}
    \country{USA}}
    \email{nasireisty@boisestate.edu}

@@

% \received{30 May 2024}
% \received[revised]{TBD}
\received[accepted]{22 July 2024}

@@

The 1995 movie Hackers prescient predictions regarding the ease
of breaking into computing systems in cyberspace have come to fruition. The White House Office of
the National Cyber Director (ONCD) released a report calling for the technical community to
proactively reduce the attack surface in cyberspace with a two-pronged approach~\cite{United_States_Gov2024-pp}. First, we need to
address the root cause of many of the most heinous cyber attacks: memory-unsafe programming
languages. Second, we need to establish better cybersecurity quality
metrics so we can have a better understanding of the cybersecurity landscape.

@@

\subsection{Research Questions}
\label{sec:researchQuestions}

We have defined the following research questions:

\begin{itemize}
    \item \textbf{RQ1: What are the existing approaches for implementing operating system kernels in Rust?}

    Although Rust in the Linux kernel is recent, we want to see what approaches researchers have explored in any operating system. Experiential and research kernels can provide new ideas that can be integrated into the Linux kernel.
    \item \textbf{RQ2: What are the performance implications of using Rust for operating system kernel development?}

    The C programming language has the advantage of 40 years of improvements and lessons learned regarding how to write fast code. Rust, being a new language, still has many unexplored areas regarding how to write fast code, especially in kernel space. Therefore, we want to explore the performance implications of Rust for kernel development in terms of throughput, latency, and resource utilization.
    \item \textbf{RQ3: What are the major challenges and limitations when developing operating system kernels in Rust?}% particularly in comparison to other languages?

      Kernel space presents unique challenges for the software engineer using Rust. Rust was not designed to implement kernels and is more restrictive than C. Thus, we want to identify if there are any stumbling blocks that have already been found when using Rust in kernel development.
    \item \textbf{RQ4: What are the lessons learned when developing operating systems kernels in Rust?}

    Besides technical challenges and limitations, developers' perceptions are very important in software development. Therefore, through this research question, we want to find developers' reflections on what went well and what didn't go well while developing kernels in Rust.
\end{itemize}

@@

We employed a robust multi-step search strategy across four databases: ACM Digital Library,
IEEE Xplore, Scopus, and Springer Link to find all the current research. 
%These databases cover a wide range of current software engineering and computer science literature. 
We leveraged the advanced search features of all three databases to search the titles and abstracts for
keywords using boolean search operators.

@@

We searched all four databases with the keywords listed in table~\ref{tab:keywords} between January
1, 2019 and April 1, 2024. We choose to look at only the previous 5 years of research in order to capture the
bleeding-edge research that is currently being done. While both ACM and IEEE had mutually exclusive results, Scopus had 44
duplicates that needed to be removed and Springer Link didn't have any new papers that were not also in ACM or IEEE. We then used Google Scholar to do forward and backward snowballing.

\begin{table}
\begin{tabular}{|| p{2cm}| p{4cm} | p{1cm} ||}
 \hline
 Database & Query & Results \\
 \hline\hline
 ACM  & (Abstract:("operating system" kernel linux OS) AND Abstract:(rust) )OR (Title:("operating
 system" OS kernel linux) AND Title:(rust))  & 30 \\
 IEEE & ("All Metadata":"operating system" AND "All Metadata":rust) OR ("All Metadata":kernel AND
 "All Metadata":rust) OR ("All Metadata":linux AND "All Metadata":rust) & 39 \\
 Scopus & TITLE-ABS-KEY ( ( "operating system" OR kernel OR linux ) AND rust ) AND PUBYEAR > 2018
 AND PUBYEAR < 2025 AND ( LIMIT-TO ( SUBJAREA , "COMP" ) ) AND ( LIMIT-TO ( LANGUAGE , "English" ) )
 & 87 \\
 Springer Link & Operating System Rust Kernel (Conference Paper, Article, Research article) Subdiscipline: Software engineering/programming and operating systems & 29 \\
 \hline
\end{tabular}
\caption{Search Queries used for each Database.}
\label{tab:keywords}
\end{table}

@@

Following the guidelines outlined by Kitchenham and Charters~\cite{Stuart2007-cc} we set the
inclusion and exclusion criteria based on our research questions outlined in
section~\ref{sec:researchQuestions}. We only considered papers that are written in English and published
in conferences, journals, and workshops. The published papers should
describe using the Rust programming language for either developing a new kernel, extending an
existing kernel, or authoring drivers. We included any type of kernel architecture, including
monolithic kernel, microkernel, or unikernel in both the embedded and non-embedded space, as long as
the paper is using the Rust programming language in some way. Papers that describe solutions that
reside 100\% in user space or papers that relied too heavily on a component written in C were excluded. Early efforts to use Rust in kernel development
were impacted by changes and updates in both the Rust language and compiler and relied on unstable
features, which were thus excluded from this review.

@@

Once we had our initial collection of papers from the databases, we merged all duplicate papers into
one record and then started our first pass, which consisted of reading the title and abstract and
either marking the paper as \textit{include} or \textit{exclude}. Once we completed the initial screening, we were left with 26 papers that needed an in-depth reading and analysis. We found an additional 6 papers during forward and backward snowballing that were also reviewed, bringing the total number of papers to 32. The first author manually reviewed all 32 papers, built the first dataset, and conducted the analysis. The second author then verified all the data and results. Since the results were straightforward and came from a comparatively small number of papers, there were no discrepancies or missing results to solve.

@@

For the first approach, Boos et al.~\cite{Boos2020-zh} created an experimental operating system
named Theseus that operates in a single address space and single privilege level and uses properties
of the Rust programming language to realize isolation instead of relying on hardware. This novel
approach pushes some of the roles that a traditional operating system would take onto the
compiler. The design of Theseus uses a novel cell-based structure where ownership of memory and
resources is enforced by the compiler, thus avoiding these checks during runtime to get improved
performance. In addition to memory safety, Theseus has a goal of addressing the state spill
problem. State spill happens when a single service in an operating system can harbor a state change
induced by interacting with other services. That state change can eventually cause a system crash or
system instability at some point in the future, making it very difficult to track down the problem
due to the separation between when the problem occurred and when it was observed. Theseus OS is an
operating system that was designed to match the Rust language instead of the more traditional route
of matching the hardware. Li and Sato~\cite{Li2024-yb} explored using Rust to write an exokernel named W-Kernel. The authors proposed a novel architecture for an OS by embedding a WebAssembly (WASM) runtime into the kernel that can run programs written in any language that can compile down to WASM.

For the second approach, researchers take a more traditional micro-kernel implementation~\cite{Chen2023-wb, Liang2021-bo}. By pushing most kernel functionality
to user space, the authors limited the surface area that needs to be proved. Chen X et al.~\cite{Chen2023-wb} took
their implementation one step further by formally verifying their new micro-kernel named Atmosphere
by using both the liner type system of Rust in combination with an SMT solver. They were
able to get a 7.5:1 proof-to-code ratio, which is higher than other formerly verified micro-kernels
SeL4 and CeriKOS both have proof code ratios of 19:1 and 20:1, respectively. While this
approach was novel drivers running in userspace are still not formally verified and don't have to be
written in Rust.

Finally, researchers have taken existing Unikernels written in C and ported them to Rust. Unikernels are designed to do one thing
and one thing only, thus limiting the complexity that a general-purpose operating system has to
deal with~\cite{Madhavapeddy2014-zw}. This reduced complexity allowed  Lankes et al.~\cite{Lankes2019-cm} to take HermitCore and rewrite it in Rust. The new
Unikernel named RustyHermit consists of only 3.27\% unsafe Rust, with the rest of the code base consisting of safe Rust, dramatically decreasing the surface area where memory errors could originate.

@@

\begin{table}
    \begin{tabular}{||l|l|l||}
    \hline
    No. & Implication & Studies that Reported the challenge\\
    \hline\hline
    1 & Performance & \cite{Gonzalez2023-ek}, \cite{Li2024-be}, \cite{Ma2023-ef}\\
    2 & Throughput & \cite{Gonzalez2023-ek}\\
    3 & Latency & \cite{Culic2022-bk} \\
    \hline
  \end{tabular}
  \caption{Performance Implications of Rust in the Kernel}
    \label{tab:RQ2}
\end{table}

@@

Culic et al.~\cite{Culic2022-bk} looked at latency issues in Tock. Tock is a new operating
system written in Rust that is designed to run on embedded systems but does not provide real-time capabilities. The
authors attempted to add real-time capabilities by integrating eBPF into the Tock kernel to improve
the interrupt handler's response time. The authors found that early work (still in the prototype
stage) lowers the response times of the system and interrupts response times 3x.

@@

Li et al.~\cite{Li2024-be} explored the feasibility of using Rust in kernel space. The authors took
an existing component, the Out of Memory (OOM), and implemented a replacement using the Rust
programming language. The non-encapsulated interface Rust component which was almost identical to
the original C component, only introduced a 0.7\% overhead. The encapsulated Rust component, on the
other hand, added a 3\% performance overhead.

@@

\begin{table*}
    \begin{tabular}{p{1cm} p{5cm} p{10cm}}
        \hline
        No. & Challenge & Description\\
        \hline
        1 & Binary Size~\cite{Ayers2022-sf}  &
        \begin{itemize}
            \item Deeply ingrained monomorphization which increased the size of the rust binaries
            \item Compiler optimizations are not as mature as some C based compilers, thus increasing binary size
            \item Hidden data structures and data
            \item Sub-optimal compiler generated support code
        \end{itemize}
        \\
        \hline
        2 & Missing Features~\cite{Burtsev2021-mh} &
        \begin{itemize}
            \item Supporting trait bounds on functions and closures with any number of arguments
            \item Expose type information in procedural macros
            \item Support a collision-free, unique type identifier

            \item Support typed assembly language for Rust
            \item Support trusted build environments
            \item Provide software-only stack guard with extensible probing interface
            \item Develop zero-copy serialization of “plain-old” data structures
        \end{itemize}
        \\
        \hline
        3 & Soundness~\cite{Klimt2023-ob} &
        \begin{itemize}
            \item Unsynchronized Global State - any use of mutable statics is unsafe
            \item C-Style Abstractions - Use rust style abstractions to properly encapsulate internal unsafe usage of raw pointers
            \item Aliasing Mutable References - Giving out raw pointers to memory that is also referenced mutably
            \item Re-implementing Memory Access - instead of accessing specific memory regions in assembly, creating and using references to the whole region is preferable
        \end{itemize}
        \\
        \hline
        4 & Panics \cite{Ma2023-ef},\cite{Burtsev2021-mh} &
        \begin{itemize}
            \item Support extendable, no\_std unwind library
            \item Stack unwinding in embedded environments
        \end{itemize}
        \\
        \hline
        5 & C Interop \cite{Miller2019-xm}, \cite{Li2021-xo} &
        \begin{itemize}
            \item Kernel interfaces, while designed for extensibility, are not designed for type safety
            \item Hybrid Code Flow. The Rust compiler can not track ownership when switching between modules written in C and Rust
        \end{itemize}
        \\
    \hline
  \end{tabular}
  \caption{Challenges Unique to the Rust Programming Language}
    \label{tab:RQ3}
\end{table*}

@@

A primary concern for programming languages that are used for operating system development is the
size of the binary especially in an embedded environment, we must
be careful to keep the size of the Rust component as close as we can to the C component so we can
still run on the same hardware. Li et al.~\cite{Li2024-be} found that there was only a 0.06\% size increase when compared to the original C implementation.  Ayers et al.~\cite{Ayers2022-sf} focused
on reducing the size of binaries produced by the rust compiler while working with Tock OS. They were
able to identify several  causes of binary growth that are specific to the Rust programming language
and have the following 5 recommendations when using Rust in a size-constrained environment:
\begin{itemize}
  \item Minimize Length + Instantiations of Generic Code
  \item Use Trait Objects Sparingly
  \item Don't Panic
  \item Carefully Use Compiler Generated Support Code
  \item Don't use static mut
\end{itemize}

@@

Klimt et al.~\cite{Klimt2023-ob} details the lessons learned and challenges when implementing
Theseus. First, they found that it is impossible to write a complete operating system in 100\% safe
Rust. For example, when writing a memory management system, raw pointers must be used to
modify the hardware. Despite the limitations imposed by the hardware, the authors describe how
Theseus leveraged intralingual design to maximize the compiler's role in enforcing correctness. By
leveraging Rust's type system and borrow checker memory safety, correct ownership transfer can be
achieved at a higher level than what could be done in C. The authors also detailed some of the
limitations of intralingual design, such as not being as expressive as many other formal verification
techniques due to the limited invariance that the type system can enforce. One of the most
important lessons learned was the insight that a linear type system itself cannot guarantee
the uniqueness of the resource represented, such as when a memory resource may overlap. The authors
introduce the idea of using a hybrid approach of verification where they leverage both the linear
type system and an SMT solver. The authors also explore the bootstrapping problem with rust
systems. In operating systems written in C, the kernel provides the ownership root to applications running on
top. Finding a new ownership root is an open research question with regard to using Rust for
operating system development.

@@

\textit{\textbf{Internal Threats.}} 
%Internal validity refers to the degree to which the review accurately establishes
%causal relationships.
Potential internal threats include:

\begin{itemize}
  %\item Publication Bias: Studies demonstrating positive outcomes of using Rust for kernel
    %development may be more likely to be published and cited, leading to an overestimation of its
    %benefits.
  \item \textit{Quality of Studies:} The methodological rigor of the included studies varies, and some may
    suffer from design flaws or biases that were not adequately controlled for, thus affecting the
    reliability of their findings.
  \item  \textit{Selection Criteria:} The inclusion and exclusion criteria for selecting relevant studies might inadvertently bias the review towards certain types of research, such as those reporting successful integration while underrepresenting studies detailing challenges and failures.
\end{itemize}

\textit{\textbf{External Threats.}} 
%External validity pertains to the generalizability of the review's findings
%beyond the specific context of the reviewed studies. 
Threats to external validity include:

\begin{itemize}
  \item \textit{Context-Specific Findings:} Many studies focus on specific kernel modules or use-cases, which
    may not be representative of the broader kernel environment. The success of Rust in isolated
    components does not necessarily translate to the entire kernel.
  \item \textit{Temporal Changes:} The rapidly evolving nature of both the Rust programming language and the
    Linux kernel means that findings from older studies may no longer be applicable, as improvements
    and changes in both domains can alter the landscape significantly.
    \item \textit{Definition Ambiguity:} The term ``Kernel development in Rust'' encompasses a wide range of activities,
    from minor module development to complete subsystem rewrites. Variations in how researchers and
    developers interpret this integration could lead to inconsistent findings.
\end{itemize}

\textbf{\textit{Conclusion Threats}.}
%Conclusion validity involves the degree to which the conclusions drawn from the
%review are credible and supported by the data.
Potential conclusion threats include:

\begin{itemize}
  \item \textit{Heterogeneity of Studies:} The diverse methodologies, metrics, and contexts of the included studies
    can lead to challenges in synthesizing findings and drawing unified conclusions. For example, it
    is very difficult to directly compare a micro-kernel, Unikernel, and hybrid Kernel.
  \item \textit{Reviewer Bias:} Personal biases of the reviewers in interpreting data and making judgments
    about study quality and relevance may skew the results
\end{itemize}

We made efforts to minimize this through predefined criteria and multiple reviewers, but some
degree of subjectivity is inevitable.  In addressing these threats, we employed rigorous methods for
study selection, data extraction, and analysis, and we remain transparent about the
limitations. 
%Future studies could mitigate these threats further by incorporating a broader range of sources, using more precise definitions, and employing meta-analytic techniques to quantitatively synthesize findings. 
Despite these threats, our results provide a valuable synthesis of the current
state of research on the use of Rust in kernel development, highlighting both its potential benefits
and the challenges that need to be addressed.
